---
title: "Interplay between glycolytic and acetate metabolisms in <i>Escherichia coli</i><br></br>"
output:
  html_document:
    df_print: paged
    toc: yes
---

 

This notebook performs computational analyses detailed in:

> From toxic waste to beneficial nutrient: acetate boosts *Escherichia coli* growth at low glycolytic flux.
>
> P. Millard, T. Gosselin-Monplaisir, S. Uttenweiler-Joseph, B. Enjalbert
>
> bioRxiv preprint, doi: [10.1101/2022.09.20.506926](https://doi.org/10.1101/2022.09.20.506926)

This R Markdown notebook is available at <https://github.com/MetaSys-LISBP/glucose_acetate_interplay>

Author: Pierre Millard ([pierre.millard\@insa-toulouse.fr](mailto:pierre.millard@insa-toulouse.fr){.email})

Copyright INRAE, 2023

## 1. Initialize environment

Load libraries and initialize environment.

```{r}
source("misc.R")
```

## 2. Simulate response of *E. coli* to changes of glycolytic activity and acetate concentration

### 2.1. Run simulations

Load the kinetic model detailed at doi: 10.7554/eLife.63661 and available from Biomodels database ([https://www.ebi.ac.uk/biomodels/)](https://www.ebi.ac.uk/biomodels/)) with the identifier MODEL2005050001.

```{r}
loadModel("model/Millard2020_Ecoli_glc_ace_kinetic_model.cps")
```

Delete events and fix concentrations of acetate, glucose and biomass.

```{r}
deleteEvent(getEvents()$key)
setSpecies(key="Ace_out", type="fixed")
setSpecies(key="Glc", type="fixed")
setSpecies(key="X", type="fixed")
```

Define the range of acetate concentration and glycolytic activity used to run simulations.

```{r}
#   number of steps
n_step <- 50

#   acetate concentration (between 0.1 and 100 mM)
ace_concentration <- 10**seq(-1, 2, length.out = n_step)

#   glycolytic activity (between 20 and 100 % of initial Vmax)
glc_vmax <- getParameters(key="(glc_upt).V")$value*seq(0.2*n_step, n_step)/n_step
```

Calculate steady-state fluxes for each acetate concentration and level of glycolytic activity.

```{r}
# create empty matrix to store simulation results
fluxes <- c("Values[v_growth_rate]", "Values[v_glc_uptake]", "Values[v_ace_net]")
simulation_results <- array(NA, dim=c(n_step*0.8+1, n_step, length(fluxes)+1), dimnames=list(inhibition=NULL, ace_concentration=NULL, specie=c("ace_concentration", fluxes)))

# run simulations

# for each glycolytic activity
for (k in seq(n_step*0.8+1)){
  
  # create empty matrix to store simulation results for a given glycolytic activity (k)
  res_ace_range <- matrix(NA, nrow=n_step, ncol=length(fluxes)+1, dimnames=list(r=NULL, c=c("ace_concentration", fluxes)))
  
  # update Vmax_glycolysis
  setParameters(key="(glc_upt).V", value = glc_vmax[k])
  applyInitialState()
  
  # for each acetate concentration
  for (j in seq(n_step)){
    
    # update acetate concentration
    setSpecies(key="Ace_out", initial_concentration=ace_concentration[j])
    applyInitialState()
    
    # calculate steady-state fluxes
    res_ss <- runSteadyState()
    
    # save steady-state fluxes for glycolytic activity k & acetate concentration j
    res_ace_range[j,] <- c(ace_concentration[j], unlist(res_ss$global_quantities[res_ss$global_quantities$key %in% fluxes, "value"]))
  }
  
  # save steady-state fluxes for all acetate concentrations at glycolytic activity k
  simulation_results[k,,] <- res_ace_range
  
}
```

### 2.2. Glycolytic flux

Save results as Source data.

```{r}
data_to_save <- simulation_results[,,"Values[v_glc_uptake]"]
colnames(data_to_save) <- ace_concentration
rownames(data_to_save) <- glc_vmax*0.00177
write.csv(data_to_save, "results/SourceData/Figure 1/1B/growth_rate.csv", quote=FALSE)
```

Plot.

```{r Figure_1B, fig.width = 6.3, fig.height = 4.5, fig.path='results/Figures/', dev=c('png', 'pdf')}
filled.contour(x=log10(ace_concentration), y=(glc_vmax/max(glc_vmax))*100, z = t(simulation_results[,,3]), nlevels = 10,
                 plot.axes = { axis(1, at=c(-1, 0, 1, 2), label=c(0.1, 1, 10, 100));
                 axis(2)})
```

### 2.3. Acetate flux

Save results as Source data.

```{r}
data_to_save <- simulation_results[,,"Values[v_ace_net]"]
colnames(data_to_save) <- ace_concentration
rownames(data_to_save) <- glc_vmax*0.00177
write.csv(data_to_save, "results/SourceData/Figure 1/1C/acetate_flux.csv", quote=FALSE)
```

Plot.

```{r Figure_1C, fig.width = 6.3, fig.height = 4.5, fig.path='results/Figures/', dev=c('png', 'pdf')}
cols = colorRampPalette(c('royalblue4','lightblue'))
  color.palette = function(n) hcl.colors(n, "YlOrRd", rev = TRUE)

  filled.contour(x=log10(ace_concentration), y=(glc_vmax/max(glc_vmax))*100, z = t(simulation_results[,,4]), nlevels = 10, 
                 col = c(cols(5), color.palette(4)),
                 plot.axes = { axis(1, at=c(-1, 0, 1, 2), label=c(0.1, 1, 10, 100));
                 axis(2)})
```

### 2.4. Growth rate

Save results as Source data.

```{r}
data_to_save <- simulation_results[,,"Values[v_growth_rate]"]
colnames(data_to_save) <- ace_concentration
rownames(data_to_save) <- glc_vmax*0.00177
write.csv(data_to_save, "results/SourceData/Figure 1/1D/growth_rate.csv", quote=FALSE)
```

Plot.

```{r Figure_1D, fig.width = 6.3, fig.height = 4.5, fig.path='results/Figures/', dev=c('png', 'pdf')}
filled.contour(x=log10(ace_concentration), y=(glc_vmax/max(glc_vmax))*100, z = t(simulation_results[,,2]), nlevels = 10,
                 plot.axes = { axis(1, at=c(-1, 0, 1, 2), label=c(0.1, 1, 10, 100));
                 axis(2)})
```

## 3. Run sensitivity analysis

To estimate the precision on model predictions, we first run a global sensitivity analysis to estimate precision on model parameters. We apply Monte Carlo analysis to generate 500 sets of parameters by fitting noisy experimental data taken from Millard et al. (2021).

```{r}
loadModel("model/Millard2020_Ecoli_glc_ace_model_4_MC.cps")

# create calibration data files for monte carlo simulations
file.copy("model/data_1mM.txt", "model/data_1mM_MC.txt", overwrite=TRUE)
file.copy("model/data_10mM.txt", "model/data_10mM_MC.txt", overwrite=TRUE)
file.copy("model/data_30mM.txt", "model/data_30mM_MC.txt", overwrite=TRUE)

# number of Monte Carlo iterations
n_iter_mc <- 5

# get parameters for the best fit
res_PE_model4 <- runParameterEstimation(method = "Statistics", model = getCurrentModel())

# list used to set up global sensitivity analysis and store calculation results
fit_results <- list("1" = list("initial_concentrations"= c("Ace0_out"=0.92, "Glc"=12.9, "X"=0.0671),
                               "species_idx" = c("[Ace0_out]_0"=1, "[Glc]_0"=1, "[X]_0"=1),
                               "data_exp" = read.table("model/data_1mM.txt", header=TRUE, sep="\t"),
                               "sd" = c("Ace"=0.2, "Glc"=0.5, "X"=0.045, "Ace_enr"=0.03),
                               "mapping" = c("Ace"="Acet_out", "Glc"="Glc", "X"="X", "Ace_enr"="Values[Ace_enr]")),
                    "10" = list("initial_concentrations"= c("Ace0_out"=9, "Glc"=14.192, "X"=0.07),
                                "species_idx" = c("[Ace0_out]_0"=2, "[Glc]_0"=2, "[X]_0"=2),
                                "data_exp" = read.table("model/data_10mM.txt", header=TRUE, sep="\t"),
                                "sd" = c("Ace"=0.2, "Glc"=0.5, "X"=0.045, "Ace_enr"=0.03),
                                "mapping" = c("Ace"="Acet_out", "Glc"="Glc", "X"="X", "Ace_enr"="Values[Ace_enr]")),
                    "30" = list("initial_concentrations"= c("Ace0_out"=30.2448, "Glc"=12.5, "X"=0.07227),
                                "species_idx" = c("[Ace0_out]_0"=3, "[Glc]_0"=3, "[X]_0"=3),
                                "data_exp" = read.table("model/data_30mM.txt", header=TRUE, sep="\t"),
                                "sd" = c("Ace"=0.2, "Glc"=0.5, "X"=0.045, "Ace_enr"=0.03),
                                "mapping" = c("Ace"="Acet_out", "Glc"="Glc", "X"="X", "Ace_enr"="Values[Ace_enr]")),
                    "res_par" = res_PE_model4$parameters[, c("parameter", "value")],
                    "res_cost" = c(res_PE_model4$main$objective_value),
                    "res_converged" = c(test_khi2(nb_points=152, k_val=res_PE_model4$main$objective_value, nb_par=11)$`p-value, i.e. P(X^2<=value)` < 0.95))
```

Simulate data from the best fit.

```{r}
# simulate data for the best fit
for (i in c("1", "10", "30")){
  # update initial conditions (Glc, Ace and biomass concentrations)
  for (j in names(fit_results[[i]][["initial_concentrations"]])){
    setSpecies(key=j, initial_concentration=fit_results[[i]][["initial_concentrations"]][j])
  }
  # apply initial state
  applyInitialState()
  # run time course simulation
  tmp <- as.matrix(runTimeCourse()$result)
  # get complete simulation results
  idx_t_max <- (tmp[,"Time"] <= max(fit_results[[i]][["data_exp"]]$time))
  fit_results[[i]]$simulations <- array(NA, dim=c(n_iter_mc+1, sum(idx_t_max), ncol(tmp)), dimnames=list(iter=NULL, time=NULL, specie=colnames(tmp)))
  fit_results[[i]]$simulations[1,,] <- tmp[idx_t_max,]
  # get index of simulated data at experimental time points
  fit_results[[i]]$t_idx <- match(fit_results[[i]][["data_exp"]]$time, fit_results[[i]]$simulations[1,,"Time"])
  # get simulated data that match experimental time points
  sim_data <- fit_results[[i]][["data_exp"]]
  for (k in names(fit_results[[i]][["mapping"]])){
    sim_data[,k] <- fit_results[[i]]$simulations[1,fit_results[[i]]$t_idx,fit_results[[i]][["mapping"]][[k]]]
  }
  fit_results[[i]]$sim_mc <- sim_data
}
```

Run sensitivity analysis.

```{r eval=FALSE, include=FALSE}
# create progress bar
pb <- txtProgressBar(min=0, max=n_iter_mc, style=3)

# run global sensitivity analysis
j <- 0
while (j < n_iter_mc){
  
  # generate noisy datasets according to the experimental standard deviation
  for (i in c("1", "10", "30")){
    # initialize dataset with simulations of the best fit
    data_noise <- fit_results[[i]]$sim_mc
    nr <- nrow(data_noise)
    # add noise
    for (k in names(fit_results[[i]]$sd)){
      data_noise[,k] <- rnorm(nr, mean=data_noise[,k], sd=fit_results[[i]]$sd[[k]])
    }
    # set negative concentrations to 0
    data_noise[data_noise < 0] <- 0
    # save the noisy dataset
    write.table(data_noise, paste("model/data_", i, "mM_MC.txt", sep=""), sep="\t", row.names = FALSE, quote = FALSE)
  }
  
  # run parameter estimation (SRES followed by Hooke and Jeeves and Levenberg-Marquadt methods to refine convergence)
  # note: this protocol ensures faster convergence in comparison to PSO, which is necessary for monte carlo analysis
  res_PE_tmp <- runParameterEstimation(method = "SRES", randomize_start_values=TRUE, update_model=TRUE)
  res_PE_tmp <- runParameterEstimation(method = "HookeJeeves", randomize_start_values=FALSE, update_model=TRUE)
  res_PE_tmp <- runParameterEstimation(method = "LevenbergMarquardt", randomize_start_values=FALSE, update_model=TRUE)
  
  # check if optimization has converged, reiterate if not
  p_val_chi2 <- test_khi2(nb_points=152, k_val=res_PE_tmp$main$objective_value, nb_par=11)$`p-value, i.e. P(X^2<=value)`
  if (p_val_chi2 > 0.95){
    next
  }
  
  j = j+1
  
  # update the progress bar
  setTxtProgressBar(pb, j)
  
  # save parameter estimation results
  fit_results$res_par <- cbind(fit_results$res_par, res_PE_tmp$parameters[, "value"])
  fit_results$res_cost <- c(fit_results$res_cost, res_PE_tmp$main$objective_value)
  fit_results$res_converged <- c(fit_results$res_converged, p_val_chi2 < 0.95)
  
  # simulate all experiments
  for (i in c("1", "10", "30")){
    # update experiment-dependent parameters
    for (l in names(fit_results[[i]]$species_idx)){
      idx <- which(res_PE_tmp$parameters$parameter==l)[fit_results[[i]]$species_idx[l]]
      k <- sub(".*\\[(.*)\\].*", "\\1", l, perl=TRUE)
      setSpecies(key=k, initial_concentration=as.numeric(res_PE_tmp$parameters[idx,"value"]))
    }
    # apply initial state
    applyInitialState()
    # run time course simulations
    tmp <- as.matrix(runTimeCourse()$result)
    # get complete simulation results
    fit_results[[i]]$simulations[j+1,,] <- tmp[tmp[,"Time"] <= max(fit_results[[i]][["data_exp"]]$time),]
  }
}

# close progress bar
close(pb)
```

Delete temporary (noisy) calibration data files.

```{r include=FALSE}
file.remove("model/data_1mM_MC.txt")
file.remove("model/data_10mM_MC.txt")
file.remove("model/data_30mM_MC.txt")
```

Save global sensitivity analysis results for further use.

```{r eval=FALSE, include=FALSE}
save(fit_results, file = "results/mc_results.RData")
```

## 4. Glycolytic control of the acetate flux, in absence or presence of acetate

### 4.1. Load experimental data

```{r}
exp_data <- read.table("data/data_E_coli_K12_MG1655_Figure_2.txt", header=TRUE, sep="\t")
```

Calculate the total carbon uptake rate and the contribution of acetate to carbon uptake.

```{r}
# calculate total carbon uptake flux
Ctotal <- -exp_data$qACE*2 + exp_data$qGLC*6

# calculate the contribution of acetate to carbon uptake (qACE/(qGLC+qACE))
contribution_acetate <- -exp_data$qACE*2 / (-exp_data$qACE*2 + exp_data$qGLC*6) * 100

# add calculated metrics to measurements dataframe
exp_data <- cbind(exp_data, Ctotal, contribution_acetate)

# create mask for data obtained at low acetate concentration (< 2 mM)
filter <- (exp_data$ACE<2)
```

### 4.2. Run simulations

Simulate data at 0.1 and 30 mM acetate concentration (with sensitivity analysis).

```{r}
load(file="results/mc_results.RData")
```

```{r}
# load model
loadModel("model/Millard2020_Ecoli_glc_ace_kinetic_model_sensitivity_analysis.cps")

# remove events and fix concentrations of actate, glucose and biomass
deleteEvent(getEvents()$key)
setSpecies(key="Ace_out", type="fixed")
setSpecies(key="Glc", type="fixed")
setSpecies(key="X", type="fixed")
```

Simulate steady-state fluxes for all set of estimated parameters.

```{r eval=FALSE, include=FALSE}
# run simulations
n_glc_steps <- 50
range_glc_vmax <- seq(0.2*n_glc_steps, n_glc_steps)/n_glc_steps
glc_vmax <- getParameters(key="(glc_upt).V")$value*range_glc_vmax

ace_concentration <- c(0.1, 30)
fluxes <- c("Values[v_growth_rate]", "Values[v_glc_uptake]", "Values[v_ace_net]")

# sensitivity analysis
# run simulations for all sets of parameters
simulation_results <- array(NA, dim=c(ncol(fit_results$res_par)-1, length(glc_vmax), length(ace_concentration), length(fluxes)+2), dimnames=list(iter=NULL, glc_vmax=NULL, ace_concentration=NULL, specie=c("ace_concentration", c(fluxes, "Total_C_uptake"))))

# create progress bar
pb <- txtProgressBar(min=0, max=ncol(fit_results$res_par)-1, style=3)

for (i in seq(ncol(fit_results$res_par)-1)){
  print(i+1)
  rp <- c(fit_results$res_par[,i+1])
  names(rp) <- fit_results$res_par[,"parameter"]
  model <- update_params(getCurrentModel(), rp)
  
  res_glc_ace_range <- array(NA, dim=c(length(glc_vmax), length(ace_concentration), length(fluxes)+2), dimnames=list(glc_vmax=NULL, ace_concentration=NULL, specie=c("ace_concentration", fluxes, "Total_C_uptake")))

  
  for (g in seq(length(range_glc_vmax))){
    
    model <- update_params(getCurrentModel(), rp)
    glc_act <- getParameters(key="(glc_upt).V")$value * range_glc_vmax[g]
    setParameters(key="(glc_upt).V", value = glc_act)
    applyInitialState()
    
    res_ace_range <- matrix(NA, nrow=length(ace_concentration), ncol=length(fluxes)+2, dimnames=list(r=NULL, c=c("ace_concentration", fluxes, "Total_C_uptake")))
    for (j in seq(length(ace_concentration))){
      setSpecies(key="Ace_out", initial_concentration=ace_concentration[j])
      applyInitialState()
      res_ss <- runSteadyState()
      total_C_flux <- res_ss$global_quantities[res_ss$global_quantities$key == "Values[v_glc_uptake]", "value"]*6 - res_ss$global_quantities[res_ss$global_quantities$key == "Values[v_ace_net]", "value"]*2
      res_ace_range[j,] <- c(ace_concentration[j], unlist(res_ss$global_quantities[res_ss$global_quantities$key %in% fluxes, "value"]),unlist(total_C_flux))
    }
    
    res_glc_ace_range[g,,] <- res_ace_range
  }
  simulation_results[i,,,] <- res_glc_ace_range
    
    
  # update the progress bar
  setTxtProgressBar(pb, i)
}

# close progress bar
close(pb)
```

Save simulation results for further use.

```{r eval=FALSE, include=FALSE}
save(simulation_results, file = "results/mc_sim_1_500.RData")
```

### 4.3. Glc uptake as function of glycolytic activity

```{r}
load("results/mc_sim_1_500.RData")
```

Model predictions.

```{r Figure_2A, fig.height=3.5, fig.path='results/Figures/', fig.width=4, dev=c('png', 'pdf')}
x <- seq(20, 100, by=2)

plot_with_ci_5(simulation_results, 1, x, "Values[v_glc_uptake]", col="#1F4E79", las=1, lwd=1.2, ylim=c(0,12), xlab="Glycolytic activity (%)", ylab="qGLC (mmol/gDW/h)", yaxs="i")

plot_with_ci_5(simulation_results, 2, x, "Values[v_glc_uptake]", col="#9DC3E6", add_to_plot = TRUE, lwd=1.2)
```

Save Source data.

```{r}
data_to_save <- cbind(apply(simulation_results[,,1,"Values[v_growth_rate]"], 2, mean),
                      apply(simulation_results[,,1,"Values[v_growth_rate]"], 2, sd),
                      apply(simulation_results[,,2,"Values[v_growth_rate]"], 2, mean),
                      apply(simulation_results[,,2,"Values[v_growth_rate]"], 2, sd))

colnames(data_to_save) <- c("mean_ace_0", "sd_ace_0", "mean_ace_30", "sd_ace_30")
rownames(data_to_save) <- glc_vmax/max(glc_vmax)*100

write.csv(data_to_save, "results/SourceData/Figure 2/2A/data_2A.csv", quote=FALSE)
```

Experimental validation.

```{r Figure_2E, fig.width=4, fig.height=3.5, fig.path='results/Figures/', dev=c('png', 'pdf')}
# fit polynomial model
polyfit_no_ace <- polynomial_model(exp_data[filter,], "aMG", "qGLC", 2, lpred=seq(0,100,by=0.1))
polyfit_ace <- polynomial_model(exp_data[!filter,], "aMG", "qGLC", 2, lpred=seq(0,100,by=0.1))

# plot fit and confidence intervals
plot(polyfit_no_ace$fited_curve_x, polyfit_no_ace$fited_curve_y, ylim=c(0,12), las=1, yaxs="i", ylab="qGLC (mmol/gDW/h)", xlab="[aMG] (mM)", col="#1F4E79", lwd=1.5, type="l")
lines(polyfit_ace$fited_curve_x, polyfit_ace$fited_curve_y, col="#9DC3E6", lwd=1.5)

# NOTE: uncomment the two following lines to plot 95 % confidence intervals on the polynomial fits
#polygon(c(polyfit_no_ace$fited_curve_x, rev(polyfit_no_ace$fited_curve_x)), c(polyfit_no_ace$confidence_intervals[,1], rev(polyfit_no_ace$confidence_intervals[,2])), col=adjustcolor("#1F4E79", alpha.f=0.4), border = NA)
#polygon(c(polyfit_ace$fited_curve_x, rev(polyfit_ace$fited_curve_x)), c(polyfit_ace$confidence_intervals[,1], rev(polyfit_ace$confidence_intervals[,2])), col=adjustcolor("#9DC3E6", alpha.f=0.4), border = NA)
# plot experimental data

points(exp_data$aMG[filter], exp_data$qGLC[filter], lwd=1.0, pch=21, bg="#1F4E79", col="black", cex=1.4)
points(exp_data$aMG[!filter], exp_data$qGLC[!filter], pch=22, bg="#9DC3E6", col="black", cex=1.4, lwd=1.0)
```

Save Source data.

```{r}
data_to_save <- exp_data[,c("strain", "carbon_source", "aMG", "ACE", "qGLC", "source")]
write.csv(data_to_save, "results/SourceData/Figure 2/2E/data_2E.csv", quote=FALSE)
```

### 4.4. Acetate flux as function of glycolytic activity

Model predictions.

```{r Figure_2B, fig.width=4, fig.height=3.5, fig.path='results/Figures/', dev=c('png', 'pdf')}
x <- seq(20, 100, by=2)

plot_with_ci_5(simulation_results, 1, x, "Values[v_ace_net]", col="#548235", las=1, lwd=1.2, ylim=c(-11,5), yaxs="i", xlab="Glycolytic activity (%)", ylab="qACE (mmol/gDW/h)")
abline(h=0)

plot_with_ci_5(simulation_results, 2, x, "Values[v_ace_net]", col="#A9D18E", add_to_plot = TRUE, lwd=1.2)
```

Save Source data.

```{r}
data_to_save <- cbind(apply(simulation_results[,,1,"Values[v_ace_net]"], 2, mean),
                      apply(simulation_results[,,1,"Values[v_ace_net]"], 2, sd),
                      apply(simulation_results[,,2,"Values[v_ace_net]"], 2, mean),
                      apply(simulation_results[,,2,"Values[v_ace_net]"], 2, sd))

colnames(data_to_save) <- c("mean_ace_0", "sd_ace_0", "mean_ace_30", "sd_ace_30")
rownames(data_to_save) <- glc_vmax/max(glc_vmax)*100

write.csv(data_to_save, "results/SourceData/Figure 2/2B/data_2B.csv", quote=FALSE)
```

Experimental validation.

```{r Figure_2F, fig.width=4, fig.height=3.5, fig.path='results/Figures/', dev=c('png', 'pdf')}
# fit polynomial model
polyfit_no_ace <- polynomial_model(exp_data[filter,], "aMG", "qACE", 2, lpred=seq(0,100,by=0.1))
polyfit_ace <- polynomial_model(exp_data[!filter,], "aMG", "qACE", 2, lpred=seq(0,100,by=0.1))

# plot fit and confidence intervals
plot(polyfit_no_ace$fited_curve_x, polyfit_no_ace$fited_curve_y, ylim=c(-11,5), las=1, yaxs="i", ylab="qACE (mmol/gDW/h)", xlab="[aMG] (mM)", col="#548235", lwd=1.5, type="l")

abline(h=0)

lines(polyfit_ace$fited_curve_x, polyfit_ace$fited_curve_y, col="#A9D18E", lwd=1.5)
#polygon(c(polyfit_no_ace$fited_curve_x, rev(polyfit_no_ace$fited_curve_x)), c(polyfit_no_ace$confidence_intervals[,1], rev(polyfit_no_ace$confidence_intervals[,2])), col=adjustcolor("#548235", alpha.f=0.4), border = NA)
#polygon(c(polyfit_ace$fited_curve_x, rev(polyfit_ace$fited_curve_x)), c(polyfit_ace$confidence_intervals[,1], rev(polyfit_ace$confidence_intervals[,2])), col=adjustcolor("#A9D18E", alpha.f=0.4), border = NA)

# plot experimental data
points(exp_data$aMG[filter], exp_data$qACE[filter], lwd=1.0, pch=21, bg="#548235", col="black", cex=1.4)
points(exp_data$aMG[!filter], exp_data$qACE[!filter], pch=22, bg="#A9D18E", col="black", cex=1.4, lwd=1.0)
```

Save Source data.

```{r}
data_to_save <- exp_data[,c("strain", "carbon_source", "aMG", "ACE", "qACE", "source")]
write.csv(data_to_save, "results/SourceData/Figure 2/2F/data_2F.csv", quote=FALSE)
```

### 4.5. Growth rate as function of glycolytic activity

Model predictions.

```{r Figure_2C, fig.width=4, fig.height=3.5, fig.path='results/Figures/', dev=c('png', 'pdf')}
x <- seq(20, 100, by=2)

plot_with_ci_5(simulation_results, 1, x, "Values[v_growth_rate]", col = "#B63A2D", las = 1, lwd = 1.2, ylim = c(0, 0.7), xlab = "Glycolytic activity (%)", ylab = "Growth rate (h-1)", yaxs="i")

plot_with_ci_5( simulation_results, 2, x, "Values[v_growth_rate]", col = "#E49890", add_to_plot = TRUE, lwd = 1.2)
```

Save Source data.

```{r}
data_to_save <- cbind(apply(simulation_results[,,1,"Values[v_growth_rate]"], 2, mean),
                      apply(simulation_results[,,1,"Values[v_growth_rate]"], 2, sd),
                      apply(simulation_results[,,2,"Values[v_growth_rate]"], 2, mean),
                      apply(simulation_results[,,2,"Values[v_growth_rate]"], 2, sd))

colnames(data_to_save) <- c("mean_ace_0", "sd_ace_0", "mean_ace_30", "sd_ace_30")
rownames(data_to_save) <- glc_vmax/max(glc_vmax)*100

write.csv(data_to_save, "results/SourceData/Figure 2/2C/data_2C.csv", quote=FALSE)
```

Experimental validation.

```{r Figure_2G, fig.width=4, fig.height=3.5, fig.path='results/Figures/', dev=c('png', 'pdf')}
# fit polynomial model
polyfit_no_ace <- polynomial_model(exp_data[filter,], "aMG", "mu", 2, lpred=seq(0,100,by=0.1))
polyfit_ace <- polynomial_model(exp_data[!filter,], "aMG", "mu", 2, lpred=seq(0,100,by=0.1))

# plot fit and confidence intervals
plot(polyfit_no_ace$fited_curve_x, polyfit_no_ace$fited_curve_y, ylim=c(0,0.7), las=1, yaxs="i", ylab="growth rate (h-1)", xlab="[aMG] (mM)", col="#B63A2D", lwd=1.5, type="l")
lines(polyfit_ace$fited_curve_x, polyfit_ace$fited_curve_y, col="#E49890", lwd=1.5)
#polygon(c(polyfit_no_ace$fited_curve_x, rev(polyfit_no_ace$fited_curve_x)), c(polyfit_no_ace$confidence_intervals[,1], rev(polyfit_no_ace$confidence_intervals[,2])), col=adjustcolor("#B63A2D", alpha.f=0.4), border = NA)
#polygon(c(polyfit_ace$fited_curve_x, rev(polyfit_ace$fited_curve_x)), c(polyfit_ace$confidence_intervals[,1], rev(polyfit_ace$confidence_intervals[,2])), col=adjustcolor("#E49890", alpha.f=0.4), border = NA)

# plot experimental data
points(exp_data$aMG[filter], exp_data$mu[filter], lwd=1.0, pch=21, bg="#B63A2D", col="black", cex=1.4)
points(exp_data$aMG[!filter], exp_data$mu[!filter], pch=22, bg="#E49890", col="black", cex=1.4, lwd=1.0)
```

Save Source data.

```{r}
data_to_save <- exp_data[,c("strain", "carbon_source", "aMG", "ACE", "mu", "source")]
write.csv(data_to_save, "results/SourceData/Figure 2/2G/data_2G.csv", quote=FALSE)
```

### 4.6. Total carbon uptake

Model predictions.

```{r Figure_2D, fig.width=4, fig.height=3.5, fig.path='results/Figures/', dev=c('png', 'pdf')}
x <- seq(20, 100, by=2)

plot_with_ci_5(simulation_results, 1, x, "Total_C_uptake", col="#451E5D", las=1, lwd=1.2, ylim=c(0,60), xlab="Glycolytic activity (%)", ylab="Total C uptake (mmol/gDW/h)", yaxs="i")

plot_with_ci_5(simulation_results, 2, x, "Total_C_uptake", col="#C89CE4", add_to_plot = TRUE, lwd=1.2)
```

Save Source data.

```{r}
data_to_save <- cbind(apply(simulation_results[,,1,"Total_C_uptake"], 2, mean),
                      apply(simulation_results[,,1,"Total_C_uptake"], 2, sd),
                      apply(simulation_results[,,2,"Total_C_uptake"], 2, mean),
                      apply(simulation_results[,,2,"Total_C_uptake"], 2, sd))

colnames(data_to_save) <- c("mean_ace_0", "sd_ace_0", "mean_ace_30", "sd_ace_30")
rownames(data_to_save) <- glc_vmax/max(glc_vmax)*100

write.csv(data_to_save, "results/SourceData/Figure 2/2D/data_2D.csv", quote=FALSE)
```

Experimental validation.

```{r Figure_2H, fig.width=4, fig.height=3.5, fig.path='results/Figures/', dev=c('png', 'pdf')}
# fit polynomial model
polyfit_no_ace <- polynomial_model(exp_data[filter,], "aMG", "Ctotal", 2, lpred=seq(0,100,by=0.1))
polyfit_ace <- polynomial_model(exp_data[!filter,], "aMG", "Ctotal", 2, lpred=seq(0,100,by=0.1))

# plot fit and confidence intervals
plot(polyfit_no_ace$fited_curve_x, polyfit_no_ace$fited_curve_y, ylim=c(0, 60), las=1, yaxs="i", ylab="total carbon uptake (mmol/gDW/h)", xlab="[aMG] (mM)", col="#451E5D", lwd=1.5, type="l")
lines(polyfit_ace$fited_curve_x, polyfit_ace$fited_curve_y, col="#C89CE4", lwd=1.5)
#polygon(c(polyfit_no_ace$fited_curve_x, rev(polyfit_no_ace$fited_curve_x)), c(polyfit_no_ace$confidence_intervals[,1], rev(polyfit_no_ace$confidence_intervals[,2])), col=adjustcolor("#451E5D", alpha.f=0.4), border = NA)
#polygon(c(polyfit_ace$fited_curve_x, rev(polyfit_ace$fited_curve_x)), c(polyfit_ace$confidence_intervals[,1], rev(polyfit_ace$confidence_intervals[,2])), col=adjustcolor("#C89CE4", alpha.f=0.4), border = NA)

# plot experimental data
points(exp_data$aMG[filter], exp_data$Ctotal[filter], lwd=1.0, pch=21, bg="#451E5D", col="black", cex=1.4)
points(exp_data$aMG[!filter], exp_data$Ctotal[!filter], pch=22, bg="#C89CE4", col="black", cex=1.4, lwd=1.0)
```

Save Source data.

```{r}
data_to_save <- exp_data[,c("strain", "carbon_source", "aMG", "ACE", "Ctotal", "source")]
write.csv(data_to_save, "results/SourceData/Figure 2/2H/data_2H.csv", quote=FALSE)
```

Total carbon uptake as function of glucose uptake rate.

```{r Figure_2K, fig.width=4, fig.height=3.5, fig.path='results/Figures/', dev=c('png', 'pdf')}
# fit polynomial model
polyfit <- polynomial_model(exp_data[filter,], "qGLC", "Ctotal", 1, lpred=seq(0,12,by=0.01))

# plot fit
plot(polyfit$fited_curve_x, polyfit$fited_curve_y, ylim=c(0, 60), xlim=c(0,12), las=1, yaxs="i", ylab="total carbon uptake (mmol/gDW/h)", xlab="Glucose uptake (mmol/gDW/h)", col="#C45818", lwd=1.5, type="l")

# plot experimental data
points(exp_data$qGLC[filter], exp_data$Ctotal[filter], lwd=1.0, pch=21, bg="#C45818", col="black", cex=1.4)

polyfit <- polynomial_model(exp_data[!filter,], "qGLC", "Ctotal", 1, lpred=seq(0,15,by=0.01))

lines(polyfit$fited_curve_x, polyfit$fited_curve_y, col="#FEA500", lwd=1.5)

points(exp_data$qGLC[!filter], exp_data$Ctotal[!filter], lwd=1.0, pch=22, bg="#FEA500", col="black", cex=1.4)
```

Save Source data.

```{r}
data_to_save <- exp_data[,c("strain", "carbon_source", "aMG", "ACE", "Ctotal", "qGLC", "source")]
write.csv(data_to_save, "results/SourceData/Figure 2/2K/data_2K.csv", quote=FALSE)
```

Growth rate as function of glucose uptake rate.

```{r Figure_2L, fig.width=4, fig.height=3.5, fig.path='results/Figures/', dev=c('png', 'pdf')}
# fit polynomial model
polyfit <- polynomial_model(exp_data[filter,], "qGLC", "mu", 1, lpred=seq(0,12,by=0.01))

# plot fit
plot(polyfit$fited_curve_x, polyfit$fited_curve_y, ylim=c(0, 0.7), xlim=c(0,12), las=1, yaxs="i", ylab="mu (h-1)", xlab="Glucose uptake (mmol/gDW/h)", col="#C45818", lwd=1.5, type="l")

# plot experimental data
points(exp_data$qGLC[filter], exp_data$mu[filter], lwd=1.0, pch=21, bg="#C45818", col="black", cex=1.4)

polyfit <- polynomial_model(exp_data[!filter,], "qGLC", "mu", 1, lpred=seq(0,15,by=0.01))

lines(polyfit$fited_curve_x, polyfit$fited_curve_y, col="#FEA500", lwd=1.5)

points(exp_data$qGLC[!filter], exp_data$mu[!filter], lwd=1.0, pch=22, bg="#FEA500", col="black", cex=1.4)
```

Save Source data.

```{r}
data_to_save <- exp_data[,c("strain", "carbon_source", "aMG", "ACE", "mu", "qGLC", "source")]
write.csv(data_to_save, "results/SourceData/Figure 2/2L/data_2L.csv", quote=FALSE)
```

### 4.7. Contribution of acetate to carbon uptake

```{r Figure_2I, fig.width=4, fig.height=3.5, fig.path='results/Figures/', dev=c('png', 'pdf')}
# fit polynomial model
polyfit <- polynomial_model(exp_data, "qGLC", "qACE", 2, lpred=seq(0,10,by=0.01))

# note: to avoid a bug on windows platforms when the plotted polygon is bigger than the plot, see https://bugs.r-project.org/show_bug.cgi?id=18219
#polyfit$confidence_intervals[polyfit$confidence_intervals > 3] <- 3
#polyfit$confidence_intervals[polyfit$confidence_intervals < -11] <- -11
# plot fit and confidence intervals
plot(polyfit$fited_curve_x, polyfit$fited_curve_y, ylim=c(-11, 4), las=1, yaxs="i", ylab="qACE (mmol/gDW/h)", xlab="qGLC (mmol/gDW/h)", col="#BF9000", lwd=1.5, type="l", yaxs="i")

abline(h=0)

#polygon(c(polyfit$fited_curve_x, rev(polyfit$fited_curve_x)), c(polyfit$confidence_intervals[,1], rev(polyfit$confidence_intervals[,2])), col=adjustcolor("#BF9000", alpha.f=0.4), border = NA)
# plot experimental data
points(exp_data$qGLC[filter], exp_data$qACE[filter], lwd=1.0, pch=22, bg="#FFE699", col="black", cex=1.4)
points(exp_data$qGLC[!filter], exp_data$qACE[!filter], lwd=1.0, pch=21, bg="#BF9000", col="black", cex=1.4)
```

Save Source data.

```{r}
data_to_save <- exp_data[,c("strain", "carbon_source", "qACE", "Ctotal", "source")]
write.csv(data_to_save, "results/SourceData/Figure 2/2I/data_2I.csv", quote=FALSE)
```

```{r Figure_2J, fig.height=3.5, fig.path='results/Figures/', fig.width=4, dev=c('png', 'pdf')}
# fit polynomial model
polyfit <- polynomial_model(exp_data, "qGLC", "contribution_acetate", 2, lpred=seq(0,10,by=0.01))

# note: to avoid a bug on windows platforms when the plotted polygon is bigger than the plot, see https://bugs.r-project.org/show_bug.cgi?id=18219
#polyfit$confidence_intervals[polyfit$confidence_intervals > 60] <- 60
#polyfit$confidence_intervals[polyfit$confidence_intervals < -20] <- -20
# plot fit and confidence intervals
plot(polyfit$fited_curve_x, polyfit$fited_curve_y, ylim=c(-20, 60), las=1, yaxs="i", ylab="acetate contribution to C uptake (%)", xlab="qGLC (mmol/gDW/h)", col="#BF9000", lwd=1.5, type="l")

abline(h=0)

#polygon(c(polyfit$fited_curve_x, rev(polyfit$fited_curve_x)), c(polyfit$confidence_intervals[,1], rev(polyfit$confidence_intervals[,2])), col=adjustcolor("#BF9000", alpha.f=0.4), border = NA)

# plot experimental data
points(exp_data$qGLC[filter], exp_data$contribution_acetate[filter], lwd=1.0, pch=22, bg="#FFE699", col="black", cex=1.4)
points(exp_data$qGLC[!filter], exp_data$contribution_acetate[!filter], lwd=1.0, pch=21, bg="#BF9000", col="black", cex=1.4)
```

Save Source data.

```{r}
data_to_save <- exp_data[,c("strain", "carbon_source", "aMG", "qGLC", "contribution_acetate", "source")]
write.csv(data_to_save, "results/SourceData/Figure 2/2J/data_2J.csv", quote=FALSE)
```

## 5. Acetate concentration at which the acetate flux reverses as function of glycolytic flux

### 5.1. Run simulations

Load model.

```{r}
loadModel("model/Millard2020_Ecoli_glc_ace_kinetic_model_sensitivity_analysis.cps")

# remove events and fix concentrations of actate, glucose and biomass
deleteEvent(getEvents()$key)
setSpecies(key="Ace_out", type="fixed")
setSpecies(key="Glc", type="fixed")
setSpecies(key="X", type="fixed")
```

Run simulations.

```{r eval=FALSE, include=FALSE}
# run simulations
n_glc_steps <- 5
glc_vmax <- getParameters(key="(glc_upt).V")$value * seq(1, n_glc_steps)/n_glc_steps
n_step <- 50
ace_concentration <- 10**seq(-1, 2, length.out = n_step)
fluxes <- c("Values[v_growth_rate]", "Values[v_glc_uptake]", "Values[v_ace_net]")

# sensitivity analysis
# run simulations for all sets of parameters
res_e2 <- array(NA, dim=c(ncol(fit_results$res_par)-1, length(glc_vmax), n_step, length(fluxes)+2), dimnames=list(iter=NULL, glc_vmax=NULL, ace_concentration=NULL, specie=c("ace_concentration", fluxes, "ace_threshold")))

# create progress bar
pb <- txtProgressBar(min=0, max=ncol(fit_results$res_par)-1, style=3)

for (i in seq(ncol(fit_results$res_par)-1)){
  rp <- c(fit_results$res_par[,i+1])
  print(i+1)
  names(rp) <- fit_results$res_par[,"parameter"]
  model <- update_params(getCurrentModel(), rp)
  
  res_glc_ace_range <- array(NA, dim=c(length(glc_vmax), n_step, length(fluxes)+2), dimnames=list(glc_vmax=NULL, ace_concentration=NULL, specie=c("ace_concentration", fluxes, "ace_threshold")))
  
  for (g in seq(n_glc_steps)){
    
    setParameters(key="(glc_upt).V", value = glc_vmax[g])
    applyInitialState()
    
    res_ace_range <- matrix(NA, nrow=n_step, ncol=length(fluxes)+2, dimnames=list(r=NULL, c=c("ace_concentration", fluxes, "ace_threshold")))
    for (j in seq(n_step)){
      setSpecies(key="Ace_out", initial_concentration=ace_concentration[j])
      applyInitialState()
      res_ss <- runSteadyState()
      res_ace_range[j,] <- c(ace_concentration[j], unlist(res_ss$global_quantities[res_ss$global_quantities$key %in% fluxes, "value"]), NA)
    }
    res_ace_range[,"ace_threshold"] <- get_ace_threshold(res_ace_range[,"Values[v_ace_net]"], ace_concentration)
    res_glc_ace_range[g,,] <- res_ace_range
  }
  res_e2[i,,,] <- res_glc_ace_range
    
    
  # update the progress bar
  setTxtProgressBar(pb, i)
}

# close progress bar
close(pb)
```

Save simulation results for further use.

```{r eval=FALSE, include=FALSE}
save(res_e2, file = "results/mc_sim_2_500.RData")
```

### 5.2. Acetate concentration threshold

Model predictions.

```{r}
load("results/mc_sim_2_500.RData")
```

```{r Figure_3A, fig.height=3.5, fig.path='results/Figures/', fig.width=4, dev=c('png', 'pdf')}
xlab_main <- c(0.1, 1, 10, 100)
xlab_sec <- c(seq(0.2, 0.9, by=0.1), seq(2, 9, by=1), seq(20, 90, by=10))

cols <- sequential_hcl(6, "blues3")

plot_with_ci_4(res_e2[,1,,], "ace_concentration", "Values[v_ace_net]", col=cols[1], xaxt='n', las=1, lwd=1.5, xlim=c(0.1,100), ylim=c(-5,5), log="x", xlab="[acetate] (mM)", ylab="growth rate (h-1)")
abline(h=0)
for (i in seq(1, 5)){
  plot_with_ci_4(res_e2[,i,,], "ace_concentration", "Values[v_ace_net]", col=cols[i], add_to_plot=TRUE, lwd=1.5)

}
axis(side = 1, at = xlab_main, labels = TRUE)
axis(side = 1, at = xlab_sec, labels = FALSE, tcl=-0.3)
```

Save Source data.

```{r}
data_to_save <- apply(res_e2[,1,,"ace_concentration"], 2, mean)
for (i in seq(5)){
  data_to_save <- cbind(data_to_save,
                        apply(res_e2[,i,,"Values[v_ace_net]"], 2, mean),
                        apply(res_e2[,i,,"Values[v_ace_net]"], 2, sd))
}
colnames(data_to_save) <- c("ace_concentration",
                            "qACE_mean_glc=20%", "qACE_sd_glc=20%",
                            "qACE_mean_glc=40%", "qACE_sd_glc=40%",
                            "qACE_mean_glc=60%", "qACE_sd_glc=60%",
                            "qACE_mean_glc=80%", "qACE_sd_glc=80%",
                            "qACE_mean_glc=100%", "qACE_sd_glc=100%")
write.csv(data_to_save, "results/SourceData/Figure 3/3A/data_3A.csv", quote=FALSE)
```

```{r Figure_3B, fig.height=3.5, fig.path='results/Figures/', fig.width=4, dev=c('png', 'pdf')}

n_glc_steps <- 5
x <- seq(1, n_glc_steps)/n_glc_steps*100

plot(x, rev(apply(res_e2[,,1,"ace_threshold"], 2, mean)), type="l", axes=FALSE, col="#548235", las=1, lwd=1.2, ylim=c(0,16), xlab="Glycolytic activity (%)", ylab="[Ace]reversal (mM)")

axis(1, at = seq(20,100,by=20), labels = paste(seq(100,20, by=-20)), tick = TRUE)
axis(2, tick = TRUE, las=1)
box()

polygon(x=c(x, rev(x)),
          y=c(rev(apply(res_e2[,,1,"ace_threshold"], 2, mean))+rev(apply(res_e2[,,1,"ace_threshold"], 2, sd)), rev(rev(apply(res_e2[,,1,"ace_threshold"], 2, mean))-rev(apply(res_e2[,,1,"ace_threshold"], 2, sd)))),
          col=paste("#548235", "55", sep=""), border=NA)
```

Save Source data.

```{r}

data_to_save <- apply(res_e2[,,1,"ace_threshold"], 2, mean)
for (i in seq(5)){
  data_to_save <- cbind(data_to_save,
                        apply(res_e2[,i,,"Values[v_ace_net]"], 2, mean),
                        apply(res_e2[,i,,"Values[v_ace_net]"], 2, sd))
}
colnames(data_to_save) <- c("ace_concentration",
                            "qACE_mean_glc=20%", "qACE_sd_glc=20%",
                            "qACE_mean_glc=40%", "qACE_sd_glc=40%",
                            "qACE_mean_glc=60%", "qACE_sd_glc=60%",
                            "qACE_mean_glc=80%", "qACE_sd_glc=80%",
                            "qACE_mean_glc=100%", "qACE_sd_glc=100%")
write.csv(data_to_save, "results/SourceData/Figure 3/3B/data_3B.csv", quote=FALSE)
```

Experimental validation.

```{r Figure_3C, fig.height=3.5, fig.path='results/Figures/', fig.width=4, dev=c('png', 'pdf')}


# load experimental data
exp_data <- read.table("data/data_E_coli_K12_MG1655_Figure_3-4.txt", header=TRUE, sep="\t")

# plot acetate flux as function of acetate concentration, at different glycolytic flux (aMG=0, 20, 40 or 100 mM)
#pdf(file="Figure_3.pdf", width = 4, height = 4)

  cols <- rev(colorRampPalette(c("#2E75B6", "#D5E3F0"))(4))
  cols <- hcl.colors(6, "Oslo")

  # empty plot
  plot(NA, NA, type="p", log="x", xaxt='n', pch=20, las=1, xlim=c(0.1,100), ylim=c(-11,3), xlab="[acetate] (mM)", ylab="acetate flux (mmol/gDW/h)")
  abline(h=0)

  # aMG = 0 mM
  filter <- (exp_data[,"aMG"]==0)
  col <- cols[2]
  x <- exp_data[filter,"ACE"]
  y <- exp_data[filter,"qACE"]

  points(x, y, bg=col, col="black", pch=21, lwd=1)
  # uncomment the following lines to run the fits, here we just plot the fitted curves
  # note: fitting is not stable, so it might have to be run several times to obtain a correct fit
  #res_fit <- fit_sigmoid(x, y, par_ini_sigmoid, lower_sigmoid, upper_sigmoid)
  #lines(seq(0.1,130,length.out=1000), sim_sigmoid(res_fit$par, seq(0.1,130,length.out=1000)), col=col, lwd=2)
  lines(seq(0.1,130,length.out=1000), sim_sigmoid(c(-0.2460494, -1.5989450, -0.9279984, 6.9209620), seq(0.1,130,length.out=1000)), col=col, lwd=2)
  
  # aMG = 20 mM
  filter <- (exp_data[,"aMG"]==20)
  col <- cols[3]
  x <- exp_data[filter,"ACE"]
  y <- exp_data[filter,"qACE"]
  
  points(x, y, bg=col, col="black", pch=22, lwd=1)
  #res_fit <- fit_sigmoid(x, y, par_ini_sigmoid, lower_sigmoid, upper_sigmoid)
  #lines(seq(0.1,130,length.out=1000), sim_sigmoid(res_fit$par, seq(0.1,130,length.out=1000)), col=col, lwd=2)
  lines(seq(0.1,130,length.out=1000), sim_sigmoid(c(-2.454911e-01, -2.840827e+00, 5.006536e+01, 1.000000e+06), seq(0.1,130,length.out=1000)), col=col, lwd=2)

  # aMG = 40 mM
  filter <- (exp_data[,"aMG"]==40)
  col <- cols[4]
  x <- exp_data[filter,"ACE"]
  y <- exp_data[filter,"qACE"]
  
  points(x, y, bg=col, col="black", pch=23, lwd=1)
  #res_fit <- fit_sigmoid(x, y, par_ini_sigmoid, lower_sigmoid, upper_sigmoid)
  #lines(seq(0.1,130,length.out=1000), sim_sigmoid(res_fit$par, seq(0.1,130,length.out=1000)), col=col, lwd=2)
  lines(seq(0.1,130,length.out=1000), sim_sigmoid(c(-0.1506843, -7.0591207, 3.5268378, 18.6152155), seq(0.1,130,length.out=1000)), col=col, lwd=2)
  
  # aMG = 100 mM
  filter <- (exp_data[,"aMG"]==100)
  col <- cols[5]
  x <- exp_data[filter,"ACE"]
  y <- exp_data[filter,"qACE"]

  points(x, y, bg=col, col="black", pch=24, lwd=1)
  #res_fit <- fit_sigmoid(x, y, par_ini_sigmoid, lower_sigmoid, upper_sigmoid)
  #lines(seq(0.1,130,length.out=1000), sim_sigmoid(res_fit$par, seq(0.1,130,length.out=1000)), col=col, lwd=2)
  lines(seq(0.1,130,length.out=1000), sim_sigmoid(c(-0.2395667, -9.8451042, 36.7387431, 64514.4364655), seq(0.1,130,length.out=1000)), col=col, lwd=2)
  
  # add axes
  xlab_main <- c(0.1, 1, 10, 100)
  xlab_sec <- c(seq(0.2, 0.9, by=0.1), seq(2, 9, by=1), seq(20, 90, by=10))
  axis(side = 1, at = xlab_main, labels = TRUE)
  axis(side = 1, at = xlab_sec, labels = FALSE, tcl=-0.3)
```

```{r Figure_3D, fig.height=3.5, fig.path='results/Figures/', fig.width=4, dev=c('png', 'pdf')}
v0 <- get_ace_threshold(sim_sigmoid(c(-0.2460494, -1.5989450, -0.9279984, 6.9209620), seq(0.1,130,length.out=1000)),
                        seq(0.1,130,length.out=1000))
v20 <- get_ace_threshold(sim_sigmoid(c(-2.454911e-01, -2.840827e+00, 5.006536e+01, 1.000000e+06), seq(0.1,130,length.out=1000)),
                        seq(0.1,130,length.out=1000))
v40 <- get_ace_threshold(sim_sigmoid(c(-0.1506843, -7.0591207, 3.5268378, 18.6152155), seq(0.1,130,length.out=1000)),
                        seq(0.1,130,length.out=1000))
v100 <- get_ace_threshold(sim_sigmoid(c(-0.2395667, -9.8451042, 36.7387431, 64514.4364655), seq(0.1,130,length.out=1000)),
                        seq(0.1,130,length.out=1000))
  
plot(c(0, 20, 40, 100), c(v0,v20,v40,v100), type="l", col="#548235", las=1, lwd=1.2, ylim=c(0,10), xlab="aMG (mM)", ylab="[Ace]reversal (mM)")
points(c(0, 20, 40, 100), c(v0,v20,v40,v100), col="#548235", lwd=1.2, pch=20)
```

## 6. Growth rate on glucose as function of aMG and acetate concentrations

Model predictions.

```{r Figure_4A, fig.height=3.5, fig.path='results/Figures/', fig.width=4, dev=c('png', 'pdf')}


xlab_main <- c(0.1, 1, 10, 100)
xlab_sec <- c(seq(0.2, 0.9, by=0.1), seq(2, 9, by=1), seq(20, 90, by=10))

# growth rate as function of acetate concentration

#cols <- c("#052547", "#2F609A", "#7199D3", "#C0D0EE")
cols <- rev(hcl.colors(6, "Oslo"))
library("colorspace")
cols = sequential_hcl(6, "blues3")


plot_with_ci_4(res_e2[,1,,], "ace_concentration", "Values[v_growth_rate]", col=cols[1], xaxt='n', las=1, lwd=2, xlim=c(0.1,100), ylim=c(0,0.8), log="x", xlab="[acetate] (mM)", ylab="growth rate (h-1)")
#for (i in seq(1, 5)){
for (i in c(5)){
  plot_with_ci_4(res_e2[,i,,], "ace_concentration", "Values[v_growth_rate]", col=cols[i-1], add_to_plot=TRUE, lwd=2)

}
axis(side = 1, at = xlab_main, labels = TRUE)
axis(side = 1, at = xlab_sec, labels = FALSE, tcl=-0.3)
```

Experimental validation.

```{r Figure_4B, fig.height=3.5, fig.path='results/Figures/', fig.width=4, dev=c('png', 'pdf')}

###############################
# IMPACT OF ACETATE ON GROWTH #
###############################
# Output:                     #
#   - Figure_4.pdf: Fig. 4    #
###############################

# load experimental data

exp_data <- read.table("data/data_E_coli_K12_MG1655_Figure_3-4.txt", header=TRUE, sep="\t")


# plot growth rate as function of acetate concentration, at high (aMG=0mM) and low (aMG=100mM) glycolytic flux

#pdf(file="Figure_4.pdf", width = 4, height = 4)

  cols <- rev(colorRampPalette(c("#2E75B6", "#D5E3F0"))(4))
  cols <- hcl.colors(6, "Oslo")

  # empty plot
  plot(NA, NA, type="p", log="x", col=col, xaxt='n', pch=20, las=1, xlim=c(0.1,100), ylim=c(0, 0.7), xlab="[acetate] (mM)", ylab="growth rate (h-1)")
  
  # aMG = 0 mM
  filter <- (exp_data[,"aMG"]==0)
  col <- cols[2]
  x <- exp_data[filter,"ACE"]
  y <- exp_data[filter,"mu"]
  
  points(x, y, bg=cols[2], col="black", pch=21, lwd=1)
  
  df <- data.frame(x=log10(x), y=y)
  df <- df[order(df$x),]
  
  model <- lm(y~poly(x,3), data = df)
  #summary(model)
  range_ace <- seq(min(df$x), max(df$x), len=100)
  predy <- predict(model, data.frame(x=range_ace))
  lines(10**range_ace, predy, type="l", col=col, lwd=2)
  
  # aMG = 100 mM
  filter <- (exp_data[,"aMG"]==100)
  col <- cols[5]
  x <- exp_data[filter,"ACE"]
  y <- exp_data[filter,"mu"]

  points(x, y, bg=col, col="black", pch=24, lwd=1)
  df=data.frame(x=log10(x), y=y)
  df <- df[order(df$x),]
  
  model = lm(y~poly(x,3), data = df)
  #summary(model)
  i = seq(min(df$x), max(df$x), len=100)       #  x-values for line
  predy = predict(model, data.frame(x=i))
  lines(10**i, predy, type="l", col=col, lwd=2)
  
  # add axes
  xlab_main <- c(0.1, 1, 10, 100)
  xlab_sec <- c(seq(0.2, 0.9, by=0.1), seq(2, 9, by=1), seq(20, 90, by=10))
  axis(side = 1, at = xlab_main, labels = TRUE)
  axis(side = 1, at = xlab_sec, labels = FALSE, tcl=-0.3)
```

## 7. Impact of acetate on growth of WT and mutant strains

### 7.1. Statistical analyses

Load experimental data.

```{r}
exp_data <- read.table("data/data_E_coli_K12_BW25113_Figure_6-7.txt", header=TRUE, sep="\t")
```

The relative change in growth rate is calculated as:

$$\Delta mu = {\mu_{ace=10} - \mu_{ace=0} \over \mu_{ace=0}}$$

For each biological replicate, we inoculated media with and without acetate from the same pre-cultivation, and calculated the change of growth rate between these two growth experiments. We carried out at least three independent biological replicate for each condition and calculated the mean ± standard deviation from these replicates. We performed Student t-tests with a null hypothesis that the ratio between the growth rates with and without acetate is 1.

```{r}
statistical_results <- array(NA, dim=c(30, 11), dimnames=list(NULL, c("carbon source", "strain", "aMG", "ace", "mu_mean", "mu_sd", "n_replicates", "p-val (vs wt, same ace)", "p-val (vs 0 mM ace, same strain)", "relative growth change", "rgc_sd")))

k <- 0
# for each carbon source
for (c_source in unique(exp_data$carbon_source)){
  # for each strain
  for (strain in unique(exp_data$genotype)){
    # for each aMG concentration
    for (aMG in rev(unique(exp_data[((exp_data$genotype == strain) & (exp_data$carbon_source == c_source)), "aMG"]))){
      # for each acetate concentration
      for (ace in unique(exp_data$ACE)){
        # perform all statistical tests (Student t-tests with two-tailed distributions)
        # to compare the different strains/conditions, as detailed below
        k <- k+1
        mu_strain <- exp_data[((exp_data$genotype == strain) & (exp_data$ACE == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG)), "mu"]
        mu_mean <- mean(mu_strain)
        mu_sd <- sd(mu_strain)
        #cat("**********************************************\n")
        #cat("Statistical analysis for the following condition:\n")
        #cat("\nstrain: ", strain, ", carbon source: ", c_source , ", [ace]=", ace, "mM, [aMG]=", aMG, "mM\n", sep="")
        #cat("**********************************************\n")
        t_test_factor_strain <- list(p.value=NA)
        if (strain != "wt"){
          n_rep_com <- Reduce(intersect, list(exp_data[((exp_data$genotype == "wt") & (exp_data$ACE == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG)), "experiment_number"],
                                              exp_data[((exp_data$genotype == strain) & (exp_data$ACE == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG)), "experiment_number"]))
          mu_wt <- exp_data[((exp_data$genotype == "wt") & (exp_data$ACE == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG) & (exp_data$experiment_number %in% n_rep_com)), "mu"]
          mu_strain <- exp_data[((exp_data$genotype == strain) & (exp_data$ACE == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG) & (exp_data$experiment_number %in% n_rep_com)), "mu"]
          if ((length(mu_wt) > 1) & (length(mu_strain) > 1)){
            t_test_factor_strain <- t.test(mu_wt, mu_strain)
          }
          
          #cat("\n\nStatistical test:", paste("\n  wt vs ", strain, ", carbon source: ", c_source , ", [ace]=", ace, "mM\n", sep=""), sep="\n")
          #cat("\nwt: ", paste(mu_wt, collapse=", "), "\n", sep="")
          #cat(strain, ": ", paste(mu_strain, collapse=", "), "\n", sep="")
          #print(t_test_factor_strain)
        }
        t_test_factor_ace <- list(p.value=NA)
        relative_growth_change <- NA
        relative_growth_change_sd <- NA
        if (ace == 10){
          mu_no_ace <- exp_data[((exp_data$genotype == strain) & (exp_data$ACE == 0) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG)), "mu"]
          mu_ace <- exp_data[((exp_data$genotype == strain) & (exp_data$ACE == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG)), "mu"]
          relative_growth_change <- mean(mu_ace/mu_no_ace)-1
          relative_growth_change_sd <- sd(mu_ace/mu_no_ace)
          # here we compared directly the relative change of growth rate induced by 10 mM acetate
          t_test_factor_ace <- t.test(mu_ace/mu_no_ace, mu=1)
          # to compare directly the growth rates in presence and absence of acetate (with paired tests, because for
          # each strain/carbon source cultivations were carried out in parallel with and without acetate), just
          # uncomment the following line (results are similar)
          # t_test_factor_ace <- t.test(mu_ace, mu_no_ace, paired=TRUE)
          #cat("\n\nStatistical test:", paste("\n  strain: ", strain, ", carbon source: ", c_source , ", [ace]=0 mM vs [ace]=10 mM\n", sep=""), sep="\n")
          #cat("relative change of growth rates: ", paste(mu_ace/mu_no_ace-1, collapse=", "), "\n", sep="")
          #print(t_test_factor_ace)
        }
        statistical_results[k, ] <- c(c_source, strain, aMG, ace, mu_mean, mu_sd, length(mu_strain), t_test_factor_strain$p.value, t_test_factor_ace$p.value, relative_growth_change, relative_growth_change_sd)
      }
    }
  }
}
```

Save results.

```{r}
write.table(statistical_results,"results/Statistical_results_glc_gly_gal.tsv", quote=FALSE, sep="\t", row.names=FALSE)
```

### 7.2. Impact of acetate & acetic acid on glucose growth of WT, 𝚫*pgi* and 𝚫*pfkA* strains

Plot data.

```{r Figure_4C, fig.height=3.5, fig.path='results/Figures/', fig.width=3, dev=c('png', 'pdf')}
mask <- (is.element(statistical_results[,"carbon source"], c("glucose"))) & (statistical_results[,"ace"] == "10") & (is.element(statistical_results[,"strain"], c("wt")))

data_to_plot <- as.numeric(statistical_results[mask, "relative growth change"])
names(data_to_plot) <- statistical_results[mask, "strain"]
data_sd_to_plot <- as.numeric(statistical_results[mask, "rgc_sd"])
ze_barplot <- barplot(data_to_plot*100, ylim=c(-10,15), beside=TRUE, legend.text=FALSE, las=1, ylab="relative growth change (%)", xaxt='n')
abline(h=0)
error.bar(ze_barplot, data_to_plot*100, data_sd_to_plot*100)
axis(side=1,line=0,at=ze_barplot,labels=c("aMG_0", "aMG_100"),mgp=c(0,0.5,0),tck=-0.02)
box()
```

```{r Figure_4D, fig.height=3.5, fig.path='results/Figures/', fig.width=3, dev=c('png', 'pdf')}
mask <- (is.element(statistical_results[,"carbon source"], c("glucose"))) & (statistical_results[,"ace"] == "10") & (is.element(statistical_results[,"strain"], c("delta_pgi", "delta_pfkA")))

data_to_plot <- as.numeric(statistical_results[mask, "relative growth change"])
names(data_to_plot) <- statistical_results[mask, "strain"]
data_sd_to_plot <- as.numeric(statistical_results[mask, "rgc_sd"])
ze_barplot <- barplot(data_to_plot*100, ylim=c(0,150), beside=TRUE, legend.text=FALSE, las=1, ylab="relative growth change (%)", xaxt='n')
abline(h=0)
error.bar(ze_barplot, data_to_plot*100, data_sd_to_plot*100)
axis(side=1,line=0,at=ze_barplot,labels=c("pfkA", "pgi"),mgp=c(0,0.5,0),tck=-0.02)
box()
```

### 7.3. Impact of acetate & acetic acid on glucose growth of WT, 𝚫*pta* and 𝚫*acs* strains

Plot data.

```{r Figure_6, fig.height=3.5, fig.path='results/Figures/', fig.width=4, dev=c('png', 'pdf')}
mask <- ((is.element(statistical_results[,"carbon source"], c("glucose"))) & (statistical_results[,"ace"] == "10")) & (statistical_results[,"aMG"] == "100") & (is.element(statistical_results[,"strain"], c("wt", "delta_pta", "delta_acs")))

data_to_plot <- as.numeric(statistical_results[mask, "relative growth change"])
names(data_to_plot) <- statistical_results[mask, "strain"]
data_sd_to_plot <- as.numeric(statistical_results[mask, "rgc_sd"])
ze_barplot <- barplot(data_to_plot*100, ylim=c(-10,15), beside=TRUE, legend.text=FALSE, las=1, ylab="relative growth change (%)", xaxt='n')
abline(h=0)
error.bar(ze_barplot, data_to_plot*100, data_sd_to_plot*100)
axis(side=1,line=0,at=ze_barplot,labels=c("aMG_100_wt","aMG_100_pta","aMG_100_acs"),mgp=c(0,0.5,0),tck=-0.02)
box()
```

### 7.4. Impact of acetate & acetic acid on growth of WT, 𝚫*pta* and 𝚫*acs* strains on glycerol and galactose

Relative change in growth rate on glycerol.

```{r Figure_7A, fig.height=3.5, fig.path='results/Figures/', fig.width=4, dev=c('png', 'pdf')}
mask <- (is.element(statistical_results[,"carbon source"], c("glycerol"))) & (statistical_results[,"ace"] == "10")

data_to_plot <- as.numeric(statistical_results[mask, "relative growth change"])
names(data_to_plot) <- statistical_results[mask, "strain"]
data_sd_to_plot <- as.numeric(statistical_results[mask, "rgc_sd"])
ze_barplot <- barplot(data_to_plot*100, ylim=c(-10,15), beside=TRUE, legend.text=FALSE, las=1, ylab="relative growth change (%)", xaxt='n')
abline(h=0)
error.bar(ze_barplot, data_to_plot*100, data_sd_to_plot*100)
axis(side=1,line=0,at=ze_barplot,labels=c("wt", "pta","acs"),mgp=c(0,0.5,0),tck=-0.02)
box()
```

Relative change in growth rate on galactose.

```{r Figure_7B, fig.height=3.5, fig.path='results/Figures/', fig.width=4, dev=c('png', 'pdf')}
mask <- (is.element(statistical_results[,"carbon source"], c("galactose"))) & (statistical_results[,"ace"] == "10")

data_to_plot <- as.numeric(statistical_results[mask, "relative growth change"])
names(data_to_plot) <- statistical_results[mask, "strain"]
data_sd_to_plot <- as.numeric(statistical_results[mask, "rgc_sd"])
ze_barplot <- barplot(data_to_plot*100, ylim=c(-20,60), beside=TRUE, legend.text=FALSE, las=1, ylab="relative growth change (%)", xaxt='n')
abline(h=0)
error.bar(ze_barplot, data_to_plot*100, data_sd_to_plot*100)
axis(side=1,line=0,at=ze_barplot,labels=c("wt", "pta","acs"),mgp=c(0,0.5,0),tck=-0.02)
box()
```

## 8. Change in pH and impact of acetic acid on growth

### 8.1. Change in pH with sodium acetate

Load experimental data.

```{r}
exp_data <- read.table("data/data_E_coli_K12_BW25113_Figure_EV1.txt", header=TRUE, sep="\t")
```

```{r}

statistical_results <- array(NA, dim=c(8, 6), dimnames=list(NULL, c("carbon source", "strain", "aMG", "ace", "delta_pH_mean", "delta_pH_sd")))

k <- 0
# for each carbon source
for (c_source in unique(exp_data$carbon_source)){
  # for each strain
  for (strain in unique(exp_data$genotype)){
    # for each aMG concentration
    for (aMG in rev(unique(exp_data[((exp_data$genotype == strain) & (exp_data$carbon_source == c_source)), "aMG"]))){
      # for each acetate concentration
      for (ace in unique(exp_data$ACE)){
        # perform all statistical tests (Student t-tests with two-tailed distributions)
        # to compare the different strains/conditions, as detailed below
        delta_pH <- exp_data[((exp_data$genotype == strain) & (exp_data$ACE == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG)), "delta_pH"]
        if (length(delta_pH) > 2){
        k <- k+1
        delta_pH_mean <- mean(delta_pH)
        delta_pH_sd <- sd(delta_pH)

        statistical_results[k, ] <- c(c_source, strain, aMG, ace, delta_pH_mean, delta_pH_sd)
          
        }
      }
    }
  }
}
```

Save results.

```{r}
write.table(statistical_results,"results/Statistical_results_glc_gly_gal_EV1.tsv", quote=FALSE, sep="\t", row.names=FALSE)
```

Plot change in pH.

```{r Figure_EV1, fig.height=4.5, fig.path='results/Figures/', fig.width=6, dev=c('png', 'pdf')}

data_to_plot <- as.numeric(statistical_results[, "delta_pH_mean"])
names(data_to_plot) <- paste(statistical_results[, "strain"], statistical_results[, "carbon source"], statistical_results[, "ace"], sep="_")
data_sd_to_plot <- as.numeric(statistical_results[, "delta_pH_sd"])
ze_barplot <- barplot(data_to_plot, ylim=c(-0.4,0.1), beside=TRUE, legend.text=FALSE, las=1, ylab="delta pH", xaxt='n')
abline(h=0)
error.bar(ze_barplot, data_to_plot, data_sd_to_plot)
axis(side=1,line=0,at=ze_barplot,labels=c("glc_aMG", "glc_aMG_ace10", "glc_aMG_ace100", "glc", "gly", "gly_ace", "gal", "gal_ace"),mgp=c(0,0.5,0),tck=-0.02, las=2)
box()
```

### 8.2. Change in pH and growth rate with acetic acid as acetate source

Load experimental data.

```{r}
exp_data <- read.table("data/data_E_coli_K12_BW25113_Figure_EV2.txt", header=TRUE, sep="\t")
```

```{r}

statistical_results <- array(NA, dim=c(8, 13), dimnames=list(NULL, c("carbon source", "strain", "aMG", "ace", "mu_mean", "mu_sd", "n_replicates", "p-val (vs wt, same ace)", "p-val (vs 0 mM ace, same strain)", "relative growth change", "rgc_sd", "delta_pH_mean", "delta_pH_sd")))

k <- 0
# for each carbon source
for (c_source in unique(exp_data$carbon_source)){
  # for each strain
  for (strain in unique(exp_data$genotype)){
    # for each aMG concentration
    for (aMG in rev(unique(exp_data[((exp_data$genotype == strain) & (exp_data$carbon_source == c_source)), "aMG"]))){
      # for each acetate concentration
      for (ace in unique(exp_data$ACETIC_ACID)){
        # perform all statistical tests (Student t-tests with two-tailed distributions)
        # to compare the different strains/conditions, as detailed below
        k <- k+1
        mu_strain <- exp_data[((exp_data$genotype == strain) & (exp_data$ACETIC_ACID == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG)), "mu"]
        mu_mean <- mean(mu_strain)
        mu_sd <- sd(mu_strain)
        
        delta_pH <- exp_data[((exp_data$genotype == strain) & (exp_data$ACETIC_ACID == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG)), "delta_pH"]
        delta_pH_mean <- mean(delta_pH)
        delta_pH_sd <- sd(delta_pH)
        #cat("**********************************************\n")
        #cat("Statistical analysis for the following condition:\n")
        #cat("\nstrain: ", strain, ", carbon source: ", c_source , ", [ace]=", ace, "mM, [aMG]=", aMG, "mM\n", sep="")
        #cat("**********************************************\n")
        t_test_factor_strain <- list(p.value=NA)
        if (strain != "wt"){
          n_rep_com <- Reduce(intersect, list(exp_data[((exp_data$genotype == "wt") & (exp_data$ACETIC_ACID == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG)), "experiment_number"],
                                              exp_data[((exp_data$genotype == strain) & (exp_data$ACETIC_ACID == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG)), "experiment_number"]))
          mu_wt <- exp_data[((exp_data$genotype == "wt") & (exp_data$ACETIC_ACID == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG) & (exp_data$experiment_number %in% n_rep_com)), "mu"]
          mu_strain <- exp_data[((exp_data$genotype == strain) & (exp_data$ACETIC_ACID == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG) & (exp_data$experiment_number %in% n_rep_com)), "mu"]
          if ((length(mu_wt) > 1) & (length(mu_strain) > 1)){
            t_test_factor_strain <- t.test(mu_wt, mu_strain)
          }
          
          #cat("\n\nStatistical test:", paste("\n  wt vs ", strain, ", carbon source: ", c_source , ", [ace]=", ace, "mM\n", sep=""), sep="\n")
          #cat("\nwt: ", paste(mu_wt, collapse=", "), "\n", sep="")
          #cat(strain, ": ", paste(mu_strain, collapse=", "), "\n", sep="")
          #print(t_test_factor_strain)
        }
        t_test_factor_ace <- list(p.value=NA)
        relative_growth_change <- NA
        relative_growth_change_sd <- NA
        if (ace == 10){
          
                    n_rep_com <- Reduce(intersect, list(exp_data[((exp_data$genotype == strain) & (exp_data$ACETIC_ACID == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG)), "experiment_number"],
                                              exp_data[((exp_data$genotype == strain) & (exp_data$ACETIC_ACID == 0) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG)), "experiment_number"]))
                    
                    
          mu_no_ace <- exp_data[((exp_data$genotype == strain) & (exp_data$ACETIC_ACID == 0) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG) & (exp_data$experiment_number %in% n_rep_com)), "mu"]
          mu_ace <- exp_data[((exp_data$genotype == strain) & (exp_data$ACETIC_ACID == ace) & (exp_data$carbon_source == c_source) & (exp_data$aMG == aMG) & (exp_data$experiment_number %in% n_rep_com)), "mu"]
          relative_growth_change <- mean(mu_ace/mu_no_ace)-1
          relative_growth_change_sd <- sd(mu_ace/mu_no_ace)
          # here we compared directly the relative change of growth rate induced by 10 mM acetate
          t_test_factor_ace <- t.test(mu_ace/mu_no_ace, mu=1)
          # to compare directly the growth rates in presence and absence of acetate (with paired tests, because for
          # each strain/carbon source cultivations were carried out in parallel with and without acetate), just
          # uncomment the following line (results are similar)
          # t_test_factor_ace <- t.test(mu_ace, mu_no_ace, paired=TRUE)
          #cat("\n\nStatistical test:", paste("\n  strain: ", strain, ", carbon source: ", c_source , ", [ace]=0 mM vs [ace]=10 mM\n", sep=""), sep="\n")
          #cat("relative change of growth rates: ", paste(mu_ace/mu_no_ace-1, collapse=", "), "\n", sep="")
          #print(t_test_factor_ace)
        }
        statistical_results[k, ] <- c(c_source, strain, aMG, ace, mu_mean, mu_sd, length(mu_strain), t_test_factor_strain$p.value, t_test_factor_ace$p.value, relative_growth_change, relative_growth_change_sd, delta_pH_mean, delta_pH_sd)
      }
    }
  }
}
```

Save results.

```{r}
write.table(statistical_results,"results/Statistical_results_glc_gly_gal_EV2.tsv", quote=FALSE, sep="\t", row.names=FALSE)
```

Plot relative change in growth rate induced by acetate.

```{r Figure_EV2A, fig.height=3.5, fig.path='results/Figures/', fig.width=3, dev=c('png', 'pdf')}
mask <- (statistical_results[,"ace"] == "10") & (statistical_results[,"carbon source"] == "glucose")

data_to_plot <- as.numeric(statistical_results[mask, "relative growth change"])
names(data_to_plot) <- statistical_results[mask, "strain"]
data_sd_to_plot <- as.numeric(statistical_results[mask, "rgc_sd"])
ze_barplot <- barplot(data_to_plot*100, ylim=c(-5,15), beside=TRUE, legend.text=FALSE, las=1, ylab="relative growth change (%)", xaxt='n')
abline(h=0)
error.bar(ze_barplot, data_to_plot*100, data_sd_to_plot*100)
axis(side=1,line=0,at=ze_barplot,labels=c("glc_aMG", "glc"),mgp=c(0,0.5,0),tck=-0.02)
box()
```

```{r Figure_EV2B, fig.height=3.5, fig.path='results/Figures/', fig.width=2, dev=c('png', 'pdf')}
mask <- (statistical_results[,"ace"] == "10") & (statistical_results[,"carbon source"] == "glycerol")

data_to_plot <- as.numeric(statistical_results[mask, "relative growth change"])
names(data_to_plot) <- statistical_results[mask, "strain"]
data_sd_to_plot <- as.numeric(statistical_results[mask, "rgc_sd"])
ze_barplot <- barplot(data_to_plot*100, ylim=c(0,10), beside=TRUE, legend.text=FALSE, las=1, ylab="relative growth change (%)", xaxt='n')
abline(h=0)
error.bar(ze_barplot, data_to_plot*100, data_sd_to_plot*100)
axis(side=1,line=0,at=ze_barplot,labels=c("gly"),mgp=c(0,0.5,0),tck=-0.02)
box()
```

```{r Figure_EV2C, fig.height=3.5, fig.path='results/Figures/', fig.width=2, dev=c('png', 'pdf')}
mask <- (statistical_results[,"ace"] == "10") & (statistical_results[,"carbon source"] == "galactose")

data_to_plot <- as.numeric(statistical_results[mask, "relative growth change"])
names(data_to_plot) <- statistical_results[mask, "strain"]
data_sd_to_plot <- as.numeric(statistical_results[mask, "rgc_sd"])
ze_barplot <- barplot(data_to_plot*100, ylim=c(0,120), beside=TRUE, legend.text=FALSE, las=1, ylab="relative growth change (%)", xaxt='n')
abline(h=0)
error.bar(ze_barplot, data_to_plot*100, data_sd_to_plot*100)
axis(side=1,line=0,at=ze_barplot,labels=c("gal"),mgp=c(0,0.5,0),tck=-0.02)
box()
```

Plot change in pH.

```{r Figure_EV2D, fig.height=3.5, fig.path='results/Figures/', fig.width=6, dev=c('png', 'pdf')}

data_to_plot <- as.numeric(statistical_results[, "delta_pH_mean"])
names(data_to_plot) <- paste(statistical_results[, "strain"], statistical_results[, "carbon source"])
data_sd_to_plot <- as.numeric(statistical_results[, "delta_pH_sd"])
ze_barplot <- barplot(data_to_plot, ylim=c(-0.4,0.1), beside=TRUE, legend.text=FALSE, las=1, ylab="change in pH", xaxt='n')
abline(h=0)
error.bar(ze_barplot, data_to_plot, data_sd_to_plot)
axis(side=1,line=0,at=ze_barplot,labels=c("glc_aMG", "glc_aMG_ace", "glc", "glc_ace", "gly", "gly_ace", "gal", "gal_ace"),mgp=c(0,0.5,0),tck=-0.02, las=2)
box()
```
